The Journey from Wild to Textbook Data: A Case Study from the National Longitudinal Survey of Youth

### 2022-02-03 paper/paper-jsdse/paper-jsdse.tex
Edited Figure 7.

Our process of cleaning builds heavily on the \texttt{tidyverse} approach (Wickham, Averick, et al. 2019). The data is first organised into ``tidy data'' (Wickham 2014) and then further wrangled using the data pipeline and split-apply-combine approach (Wickham 2011). (Tidy data shouldn't be confused with ``tame data'' which Kim, Ismay, and Chunn (2018) coined to refer to textbook data sets suitable for teaching, particularly teaching statistics.) The resulting (tame) data is provided in a new R package called \texttt{yowie} which includes the code so that the process is reproducible, and could be used to further refresh the data as new records are made available in the NLSY79 database.

@@

There are two aspects of the original data that make a direct comparison difficult. The time variable provided is ``experience in the workforce''. It is not clear how this is calculated. One would expect that there is a record of the day the individual first started a job, and this is used to adjust the year of collection. We have not been able to find this information in the database. Secondly, wages were inflation-adjusted to 1990 prices, which is not done for the refreshed data.

>>>>

Our process of cleaning builds heavily on the \texttt{tidyverse} approach (Wickham, Averick, et al. 2019). The data is first organized into ``tidy data'' (Wickham 2014) and then further wrangled using the data pipeline and split-apply-combine approach (Wickham 2011). (Tidy data shouldn't be confused with ``tame data'' which Kim, Ismay, and Chunn (2018) coined to refer to textbook data sets suitable for teaching, particularly teaching statistics.) The resulting (tame) data is provided in a new R package called \texttt{[CENSORED]} which includes the code so that the process is reproducible, and could be used to further refresh the data as new records are made available in the NLSY79 database.

@@

There are two aspects of the original data that make a direct comparison difficult. Th first one is that The time variable provided is ``experience in the workforce''. It is not clear how this is calculated. One would expect that there is a record of the day the individual first started a job, and this is used to adjust the year of collection. Secondly, wages were inflation-adjusted to 1990 prices, which is not done for the refreshed data.

\textbackslash begin\{figure\}

\{\centering \includegraphics[width=468px]{figures/compare_xp_to_yrworkforce-1}

\}

\caption{The association between the lenght of time in the workforce (years) and the work experience. The color represents the respondents (id). Generally, the longer someone were in the workforce, the higher the work experience. However, some of the respondents have much lower working experience compared their duration in the workforce.}

(\#fig:compare\_xp\_to\_yrworkforce)
\textbackslash end\{figure\}

\begin{center}\includegraphics[width=468px]{figures/compare_xp_sw-1} \end{center}

### 2022-02-03 paper/paper.Rmd
Add the experience comparison to the paper.
Edited Figure 7.

<!--AE comment: In Section 4, you talk about the variables in the original data. “One would expect that there is a record of the day the individual first started a job, and this is used to adjust the year of collection.” I’m not sure what this means. Please clarify.-->
<!--AE comment: “experience in the workforce” should have the period inside the quotations.-->
There are two aspects of the original data that make a direct comparison difficult. The time variable provided is "experience in the workforce". It is not clear how this is calculated. One would expect that there is a record of the day the individual first started a job, and this is used to adjust the year of collection. We have not been able to find this information in the database. Secondly, wages were inflation-adjusted to 1990 prices, which is not done for the refreshed data.

>>>>

<!--AE comment: In Section 4, you talk about the variables in the original data. “One would expect that there is a record of the day the individual first started a job, and this is used to adjust the year of collection.” I’m not sure what this means. Please clarify.-->
<!--AE comment: “experience in the workforce” should have the period inside the quotations.-->
There are two aspects of the original data that make a direct comparison difficult. Th first one is that The time variable provided is "experience in the workforce". It is not clear how this is calculated. One would expect that there is a record of the day the individual first started a job, and this is used to adjust the year of collection. Secondly, wages were inflation-adjusted to 1990 prices, which is not done for the refreshed data.

```{r compare_xp_to_yrworkforce, fig.cap = "The association between the lenght of time in the workforce (years) and the work experience. The color represents the respondents (id). Generally, the longer someone were in the workforce, the higher the work experience. However, some of the respondents have much lower working experience compared their duration in the workforce."}
```

```{r compare_xp_sw}
```
